---
title: "covid 19"
author: "Cyann Arevalo"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="percent Bar Chart")
```
  
  By looking at the bar chart above, we can see that although there are less males with the virus that are living, there is a higher percentage of men who have died. We can also see that there is a high percentage of women who are living with the virus but there are less women who have died than men.


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
Looking at the numerical summary, out of the total population of people who are alive, 56% of them are females while 44% of them are males. We can also see that out of the total population of people who are dead, 36% of them are females while 64% of them are males.


## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The odds of dying if you are a male is 44/1411. The odds of dying if you are a female is 25/1831. The odds ratio is (44/1411)/(25/1831). The odds of dying is 2.3 to 1.


## Conclusions

